\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Type-Preserving Mutations

### 2021-08-20 papers/icse22/main.tex
Re-structured intro and background sections

\usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{color}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{caption}
\usepackage[ruled,vlined,noend]{algorithm2e}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{tikzscale}
\usepackage{pgfplots}
\usepackage{paralist}
\usepackage{enumitem}

@@

%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
\citestyle{acmauthoryear}

@@

%% FuzzChick: PBT + code instrumentation + type-preserving mutations
%% \fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
%% for the Coq programming language that
%
Coverage-Guided, Property-Based Testing (CGPT) \cite{lampropoulos2019coverage}
is a technique that borrows ideas from the fuzzing community to generate highly
structured values while still using automatically derived generators.
%
%% Notably, this tool is implemented as an extension of \quickchick
%% \cite{denes2014quickchick}, Coq's own reimplementation of \quickcheck.
%
Instead of continuously generating random invalid test cases from scratch, CGPT
works by keeping a queue of \emph{interesting} previously executed test cases
that can be mutated to produce new ones.
%
Mutations are achieved using high-level, structure-preserving transformations
that can be derived directly from the codebase.
%
Intuitively, mutating an interesting test case in a small way (at the data
constructor level) has a higher probability of producing a new interesting test
case than generating a new one using a na\"ive generator.
%
In this light, CGPT is likely to preserve the semantic structure of the mutated
data, as mutations are applied directly at the data type level, e.g., the
program's AST in case of generating code.
%
Moreover, CGPT relies on execution traces to distinguish which executed test
cases were interesting and are therefore worth mutating --- a popular technique
known as \emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
\cite{afl}.
%
Mutated test cases are considered interesting for mutation only when they
produce new execution traces --- any other test case is simply thrown away.
%
Worth mentioning, this technique is not meant to replace sophisticated, manually
written random generators, but rather to provide an adequate testing solution
for early stages of development.
%, and while a better test suite using manually-written generators is
%still under construction.

@@

%% Validation
To validate our ideas, we use two main case studies (described in detail in
Section \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
used by \citeauthor{lampropoulos2019coverage} in their original work, and an
existing WebAssembly engine of industrial strenght written in Haskell.
%
In both case studies, we additionally compare the effect of the heuristics
implemented on top of the base testing loop of \mutagen.
%
In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
reference CGPT implementation by \citeauthor{lampropoulos2019coverage}
implemented in Coq.
%
Our results (Section \ref{sec:evaluation}) indicate that \mutagen outperforms
\fuzzchick both in terms of time to failure and failure rate.
%
On the other hand, our WebAssembly case study shows the performance of our tool
in a realistic scenario.
%
There, \mutagen is capable of reliably finding 15 planted bugs in the validator
and interpreter, as well as 3 previously unknown bugs that flew under the radar
of the existing unit test suite of this engine.
%
This case study also help us to compare the performance of our tool against a
more traditional PBT approach that does not rely on code instrumentation.
%
All in all, our evaluation encompasses more than 600 hours of computing time and
suggests that testing mutants exhaustively can be an appealing technique for
finding bugs reliably without sacrificing speed.

@@

In this section, we describe in detail the motivation, ideas and limitations
behind coverage-guided, property-based testing as depicted by
\citeauthor{lampropoulos2019coverage}.
%
To illustrate this technique, let us focus on a simple property defined over
binary trees.
%
Such a data structure can be defined in Haskell using a custom algebraic data
type with two data constructors for leaves and branches respectively:

@@

The last missing piece is a random generator of trees.
%
In \quickcheck, this is commonly done via the type class mechanism
\cite{jones1997type}, instantiating the \texttt{Arbitrary} type class with a
random generator as the implementation of the overloaded \texttt{arbitrary}
operation:

\begin{verbatim}
instance Arbitrary a => Arbitrary (Tree a) where
  arbitrary = sized gen where
  gen 0 = Leaf <$> arbitrary
  gen n = oneof [ Leaf <$> arbitrary
                , Branch <$> gen (n-1) <*> arbitrary <*> gen (n-1) ]
\end{verbatim}

\noindent This definition follows a simple type-directed generation approach,
and it is a good example of what to expect from a random generator synthesized
automatically from the codebase.
%
In esence, this generator randomly picks one data constructor from its
corresponding data type definition with uniform probability (i.e, either a
\texttt{Leaf} or a \texttt{Branch}), and then calls itself for every recursive
subexpression needed to produce a well-typed value, carefully reducing the depth
limit (controlled by \quickcheck) by a unit on each step, ensuring termination.
%
Subexpressions of any other type are generated simply by delegating the task to
their corresponding \texttt{Arbitrary} instance --- as noted by the use of
\texttt{arbitrary} when generating random payloads inside of the \texttt{Leaf}
and \texttt{Branch} constructors.

Readers familiar with RBPT will recognize that attempting to test this property
using \quickcheck directly does not work well.
%
In practice, \quickcheck will give up quite early if most of the generated test
cases get discarded because they do not satisfy the balanced precondition.
%
Sadly, most of the inputs generated by our na\"ive generator suffer from this
problem, and the interesting part of the property (\texttt{balanced (insert x t)})
is tested very sporadically as a result.
%
In other words, \emph{the precondition of our testing property becomes very
  sparse when we use simple, automatically derived generators.}

The next subsection show how coverage-guided, property-based testing can help
overcoming this limitation without having to write specialized random generators
by hand.

@@

To alleviate the problem of testing properties with non-trivial preconditions
while using automatically derived random generators,
\citeauthor{lampropoulos2019coverage} introduced \emph{coverage-guided,
  property-based testing} (CGPT), a technique that enhances the testing process
with two key characteristics:
%
\begin{inparaenum}
\item \emph{target code instrumentation}, to capture execution information from
  each test case; and
\item \emph{high-level, structure-preserving mutations}, to produce
  syntactically valid test cases by altering existing ones at the data type
  level.
\end{inparaenum}

Using code instrumentation in tandem with mutations is a well-known technique in
the fuzzing community.
%
Generic fuzzing tools like \emph{AFL} \cite{afl}, \emph{HonggFuzz}
\cite{honggfuzz}, or \emph{libFuzzer} \citeyearpar{libfuzzer} as well as
language-specific ones like Crowbar \cite{dolan2017testing} or \emph{Kelinci}
\cite{kersten2017poster} use execution traces to recognize interesting test
cases, e.g, those that exercise previously undiscovered parts of the target
code.
%
Later, such tools use generic mutators to combine and produce new test cases
from previously executed interesting ones.
%
CGPT, however, does this in a clever way.
%
Instead of mutating any previously executed test case that discovers a new part
of the code, this technique integrates these fuzzing mechanisms into the PBT
testing loop itself.

Since it is possible to distinguish semantically valid test cases from invalid
ones, i.e., those passing the testing property preconditions as opposed to those
that are discarded early, CGPT exploits this information to prefer mutating
valid test cases than discarded ones.

In addition, high-level, structure-preserving mutators are better suited for
producing syntactically valid mutants, avoiding the time wasted by using generic
low-level mutators that act over the ``serialized'' test cases and know very
little about their structure, thus producing syntactically broken mutants most
of the time.
%
This grammar-aware mutation technique has shown to be quite useful when fuzzing
systems accepting structurally complex inputs.
%
Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
specific input structure used by the system under test.
%
In CGPT, external grammars are not required.
%
The data types used by the inputs of the testing properties already describe the
structure of the random data we want to mutate in a concrete manner, and
specialized mutators acting at the data constructor level can be automatically
derived directly from their definition.

%% The next subsections describe the CGPT testing loop and structure-preserving
%% mutations in detail.

%% \subsubsection{Testing loop}

\paragraph{Testing Loop}
Outlined in Algorithm \ref{algo:fuzzchick:loop}, CGPT starts by creating two
queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
executed test cases, respectively.
%
Enqueued values are stored along with a given mutation budget, that controls how
many times a given test case can be mutated before being finally thrown away.

Once inside of the main loop, the algorithm picks the next test case using a
simple criterion: if there are valid values enqueued for mutation, it picks the
first one, mutates it and returns it, decreasing its energy by one.
%
If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
%
If none of the mutation queues contain any candidates, CGPT generates a new
value from scratch.
%
This selection process is illustrated in detail in Algorithm
\ref{algo:fuzzchick:pick}.

Having selected the next test case, the main loop proceeds to execute it,
capturing both the result (passed, discarded due to preconditions, or failed)
and its execution trace over the system under test.
%
If the test case fails, it is immediately reported as a bug.
%
If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
new path) based on its trace information and the one from previously executed
test cases (represented by \textit{TLog}).
%
If the test case does in fact discover a new path, it is enqueued at the end of
its corresponding queue, depending on whether it passed or was discarded.
%
%% (i.e., it failed the precondition).
%
This process alternates between generation and mutation until a bug is found or
we reach the test limit.

%% The energy assigned to each test case follows that of AFL's power schedule: more
%% energy to test cases that lead to shorter executions, or that discover more
%% parts of the code.
%% %
%% Moreover, to favor mutating interesting valid test cases, they get more energy
%% than those that were discarded.

\begin{algorithm}
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{FC}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\FC{P, N, gen, mut}}{
    i $\gets$ 0\;
    TLog, QSucc, QDisc $\gets$ $\varnothing$\;
    \While{i $<$ N}{
      x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
      (result, trace) $\gets$ WithTrace(P(x))\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        e $\gets$ Energy(TLog, x, trace)\;
        \eIf{\KwNot Discarded(result)}{
          Enqueue(QSucc, (x, e))\;
        }{
          Enqueue(QDisc, (x, e))\;
        }
      }
      i $\gets$ i+1\;
    }
    \KwRet Ok\;
  }
\caption{\label{algo:fuzzchick:loop}CGPT Testing Loop}
\end{algorithm}

\begin{algorithm}
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QSucc, QDisc, gen, mut}}{
    \If{\KwNot Empty(QSucc)}{
      (x,e) $\gets$ Deque(QSucc)\;
      \If{e $>$ 0}{
        PushFront(QSucc, (x, e-1))
      }
      \KwRet Sample(mut(x))\;
    }
    \uElseIf{\KwNot Empty(QDisc)}{
      (x,e) $\gets$ Deque(QDisc)\;
      \If{e $>$ 0}{
        PushFront(QDisc, (x, e-1))
      }
      \KwRet Sample(mut(x))\;
    }
    \lElse{\KwRet Sample(gen)}
  }
\caption{\label{algo:fuzzchick:pick}CGPT Seed Selection}
\end{algorithm}

%% \subsubsection{Type-preserving mutations}

\paragraph{Structure-preserving Mutators}

%% are no more than specialized random generators, parameterized
%% by the original input to be mutated.
%% %
%% They

Mutators in GCPT use a simple set of mutation operations that are randomly
applied to their inputs at the data type level.
%
In simple terms, these operations encompass
%
\begin{inparaenum}
\item \emph{swapping a data constructor}, replacing its outermost data
  constructor by another one, e.g., a \texttt{Leaf} by a \texttt{Branch} and
  vice-versa, either dropping unused subexpressions or generating new random
  ones when the new data constructor contains extra fields that cannot be filled
  using existing ones;
%% \item \emph{shrinking the value}, replacing its outermost data constructor with
%%   one that contains a subset of its fields, reusing existing subexpressions;
%% \item \emph{growing the value}, replacing its outermost data constructor with
%%   one that contains a superset of its fields, reusing existing subexpressions
%%   and generating random ones when needed;
\item \emph{returning a subexpression of the same type}, e.g, returning a left
  or right subtree when applied to a \texttt{Branch}; and
\item \emph{mutating recursively}, applying a mutation operation over an
  immediate subexpression.
\end{inparaenum}
%
We will provide more details about structure-preserving mutations in Section
\ref{sec:mutagen}.

%% \begin{figure}
%% \vspace{-5pt}
%% \begin{verbatim}
%% mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
%% mutate_tree mutate_a (Leaf x) =
%%   oneof [ do { x' <- mutate_a x; return (Leaf x')}                       -- Mutate recursively
%%         , do { l <- arbitrary; r <- arbitrary; return (Branch l x r) } ] -- Grow constructor
%% mutate_tree mutate_a (Branch l x r) =
%%   oneof [ return l                                                       -- Return subexpression
%%         , return r                                                       -- Return subexpression
%%         , return (Leaf x)                                                -- Shrink constructor
%%         , do { l' <- mutate_tree l; return (Branch l' x r) }             -- Mutate recursively
%%         , do { x' <- mutate_a    x; return (Branch l x' r) }             -- Mutate recursively
%%         , do { r' <- mutate_tree r; return (Branch l x r') } ]           -- Mutate recursively
%% \end{verbatim}
%% \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
%% \end{figure}

%% \begin{figure}
%% %% \vspace{-5pt}
%% \begin{verbatim}
%% mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
%% mutate_tree mutate_a (Leaf x) =
%%   oneof [ Leaf <$> mutate_a x
%%         , Branch <$> arbitrary <*> return x <*> arbitrary ]
%% mutate_tree mutate_a (Branch l x r) =
%%   oneof [ return l
%%         , return r
%%         , Leaf   <$> return x
%%         , Branch <$> mutate_tree l <*> return x   <*> return r
%%         , Branch <$> return l      <*> mutate_a x <*> return r
%%         , Branch <$> return l      <*> return x   <*> mutate_tree r
%% \end{verbatim}
%% \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
%% \end{figure}

%% Fig. \ref{fig:fuzzchick:mutator} illustrates a \fuzzchick-like mutator for our
%% previously used \texttt{Tree} data type example.
%% %
%% Since trees are parametric, for clarity this definition is also parameterized by
%% a mutator for the payload (\texttt{mutate\_a}), although this can be abstracted
%% away using the type class system.

%% %% (where \texttt{mutate\_tree} and
%% %% \texttt{mutate\_a} can be transparently coerced into an overloaded
%% %% \texttt{mutate} function.)

%% In this mutator, branches can be shrunk into leaves by dropping the subtrees,
%% whereas leaves can grow into branches, by reusing the payload and generating two
%% random subtrees.
%% %
%% Moreover, branches can be replaced with one of their subtrees.
%% %
%% Finally, mutations can be recursively applied over both the payload and the
%% subtrees.
%% %
%% At the top level, all these operations are put together using the \texttt{oneof}
%% combinator that randomly picks one of them with uniform probability.

@@

\item \emph{Power schedule:}
%
it is not clear how the power schedule used to assign a given budget to each
mutable test case in CGPT works in the context of high-level
structure-preserving mutations.
%
If it assigns too much energy to certain not-so-interesting seeds, some bugs
might not be discovered on a timely basis.
%
Conversely, assigning too little energy to interesting test cases might mean
that some bugs cannot be discovered at all unless the right mutation happens
within the small available energy window --- randomly generating the same test
case again later does not help, as it becomes uninteresting based on historic
trace information.

@@

%% The next two lines define the bibliography style to be used, and the
%% bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

>>>>

\usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{color}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{caption}
\usepackage[ruled,vlined,noend]{algorithm2e}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{tikzscale}
\usepackage{pgfplots}
\usepackage{paralist}
\usepackage{enumitem}
%% \usepackage[square,numbers]{natbib}

\settopmatter{printacmref=false}

@@

%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%% \citestyle{authoryear}

@@

%% FuzzChick: PBT + code instrumentation + type-preserving mutations
%% \fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
%% for the Coq programming language that
%
Coverage-Guided, Property-Based Testing (CGPT) \cite{lampropoulos2019coverage}
is a technique that borrows ideas from the fuzzing community to generate highly
structured values while still using automatically derived generators.
%
%% Notably, this tool is implemented as an extension of \quickchick
%% \cite{denes2014quickchick}, Coq's own reimplementation of \quickcheck.
%
Instead of continuously generating random invalid test cases from scratch, CGPT
works by keeping a queue of \emph{interesting} previously executed test cases
that can be mutated using high-level, structure-preserving transformations to
produce new ones.
%
%% Mutations are achieved using high-level, structure-preserving transformations
%% that can be derived directly from the codebase.
%
Intuitively, mutating an interesting test case in a small way (at the data
constructor level) has a higher probability of producing a new interesting test
case than generating a new one using a na\"ive generator.
%
In addition, high-level, structure-preserving mutators are better suited for
producing syntactically valid mutants, avoiding the time wasted by using generic
low-level mutators that act over the ``serialized'' test cases and know very
little about their structure, thus producing syntactically broken mutants most
of the time.
%
This grammar-aware mutation technique has shown to be quite useful when fuzzing
systems accepting structurally complex inputs \cite{holler2012fuzzing,
  wang2019superion, xsmith}.
%
While most of the tools following this itechnique use existing grammars to
tailor the mutators to the specific input structure used by the system under
test, external grammars are not required in CGPT.
%
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%
%% In CGPT, external grammars are not required.
%
The data types used by the inputs of the testing properties already describe the
structure of the random data we want to mutate in a concrete manner, and
specialized mutators acting at the data constructor level can be automatically
derived directly from their definition --- making strongly-typed programming
languages with expressive type systems an ideal match for CGPT.

%% In this light, CGPT is likely to preserve the semantic structure of the mutated
%% data, as mutations are applied directly at the data type level, e.g., the
%% program's AST in case of generating code.
%
Moreover, CGPT relies on execution traces to distinguish which executed test
cases were interesting and are therefore worth mutating --- a popular technique
known as \emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
\cite{afl}.
%
Mutated test cases are considered interesting for mutation only when they
produce new execution traces --- any other test case is simply thrown away.
%
Worth mentioning, this technique is not meant to replace sophisticated, manually
written random generators, but rather to provide an adequate testing solution
for early stages of development.
%, and while a better test suite using manually-written generators is
%still under construction.

@@

%% Validation
To validate our ideas, we use two main case studies (described in detail in
Section \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
used by \citeauthor{lampropoulos2019coverage} in their original work, and an
existing WebAssembly engine of industrial strenght written in Haskell.
%
In both case studies, we additionally compare the effect of the heuristics
implemented on top of the base testing loop of \mutagen.
%
In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
reference CGPT implementation developed in Coq by
\citeauthor{lampropoulos2019coverage}.
%
Our results (Section \ref{sec:evaluation}) indicate that \mutagen outperforms
\fuzzchick both in terms of time to failure and failure rate.
%
On the other hand, our WebAssembly case study shows the performance of our tool
in a realistic scenario.
%
There, \mutagen is capable of reliably finding 15 planted bugs in the validator
and interpreter, as well as 3 previously unknown bugs that flew under the radar
of the existing unit test suite of this engine.
%
This case study also help us to compare the performance of our tool against a
more traditional PBT approach that does not rely on code instrumentation.
%
All in all, our evaluation encompasses more than 600 hours of computing time and
suggests that testing mutants exhaustively can be an appealing technique for
finding bugs reliably without sacrificing speed.

@@

In this section, we briefly describe the motivation, ideas and limitations
behind CGPT as depicted by \citeauthor{lampropoulos2019coverage}.
%
To illustrate this technique, let us focus on a simple property defined over
binary trees.
%
Such a data structure can be defined in Haskell using a custom algebraic data
type with two data constructors for leaves and branches respectively:

@@

The last missing piece is a random generator of trees.
%
%% In \quickcheck, this is commonly done via the type class mechanism
%% \cite{jones1997type}, instantiating the \texttt{Arbitrary} type class with a
%% random generator as the implementation of the overloaded \texttt{arbitrary}
%% operation:
A simple implementation of such a generator (simplified to make it more
accessible%
\footnote{In practice, this code will most likely be implemented using Haskell's
  type-class mechanism, hiding the complexity behind the size limit and the payload generator.}%
) can be written as follows:

%% \begin{verbatim}
%% instance Arbitrary a => Arbitrary (Tree a) where
%%   arbitrary = sized gen where
%%   gen 0 = Leaf <$> arbitrary
%%   gen n = oneof [ Leaf <$> arbitrary
%%                 , Branch <$> gen (n-1) <*> arbitrary <*> gen (n-1) ]
%% \end{verbatim}

\begin{verbatim}
gen_tree :: Size -> Gen a -> Gen (Tree a)
gen_tree gen_a 0 = Leaf <$> gen_a
gen_tree gen_a n = oneof
  [ Leaf <$> gen_a
  , Branch <$> gen_tree (n-1) <*> gen_a <*> gen_tree (n-1) ]
\end{verbatim}

\noindent This definition follows a common type-directed generation approach,
and it is a good example of what to expect from a random generator synthesized
automatically from the codebase.
%
In esence, this generator randomly picks one data constructor from its
corresponding data type definition with uniform probability (i.e, either a
\texttt{Leaf} or a \texttt{Branch}), and then calls itself for every recursive
subexpression needed to produce a well-typed value, carefully reducing the size
limit by a unit on each step, ensuring termination by restricting itself to
generating leaves when it reaches zero.
%
Payloads are generated simply by delegating the task to an external random
generator (\texttt{gen\_a}) defined elsewhere.
%

%% Subexpressions of any other type are generated
%% simply by delegating the task to their corresponding \texttt{Arbitrary} instance
%% --- as noted by the use of \texttt{arbitrary} when generating random payloads
%% inside of the \texttt{Leaf} and \texttt{Branch} constructors.

Readers familiar with RBPT will recognize that attempting to test
\texttt{prop\_insert\_inv} using this generator in \quickcheck directly does not
work well.
%
In practice, \quickcheck will give up quite early if most of the generated test
cases get discarded because they do not satisfy the balanced precondition.
%
Sadly, most of the inputs generated by our na\"ive generator suffer from this
problem, and the interesting part of the property (\texttt{balanced (insert x t)})
is tested very sporadically as a result.
%
In other words, \emph{the precondition of our testing property becomes very
  sparse when we use simple, automatically derived generators.}

The next subsection show how GCPT helps overcoming this limitation without
having to write specialized random generators by hand.

@@

To alleviate the problem of testing properties with non-trivial preconditions
while using automatically derived random generators, CGPT
%
%% \citeauthor{lampropoulos2019coverage} introduced \emph{coverage-guided,
%%   property-based testing} (CGPT), a technique that
%
enhances the testing process with two key characteristics:
%
\begin{inparaenum}
\item \emph{target code instrumentation}, to capture execution information from
  each test case; and
\item \emph{high-level, structure-preserving mutations}, to produce
  syntactically valid test cases by altering existing ones at the data
  constructor level.
\end{inparaenum}

Using code instrumentation in tandem with mutations is a well-known technique in
the fuzzing community.
%
Generic fuzzing tools like \emph{AFL} \cite{afl}, \emph{HonggFuzz}
\cite{honggfuzz}, or \emph{libFuzzer} \citeyearpar{libfuzzer} as well as
language-specific ones like Crowbar \cite{dolan2017testing} or \emph{Kelinci}
\cite{kersten2017poster} use execution traces to recognize interesting test
cases, e.g, those that exercise previously undiscovered parts of the target
code.
%
Later, such tools use generic mutators to combine and produce new test cases
from previously executed interesting ones.
%
%% Instead of mutating any previously executed test case that discovers a new part
%% of the code, this technique integrates these fuzzing mechanisms into the PBT
%% testing loop itself.
%
In contrast with traditional fuzzing approaches, CGPT can distinguish
semantically valid test cases from invalid ones, i.e., those passing the testing
property preconditions as opposed to those that are discarded early.
%
This information is used to favour mutating valid test cases over discarded
ones.

%% Since it is possible to distinguish semantically valid test cases from invalid
%% ones, i.e., those passing the testing property preconditions as opposed to those
%% that are discarded early, CGPT exploits this information to prefer mutating
%% valid test cases than discarded ones.

The testing loop in CGPT uses two queues to store valid and discarded previously
executed test cases.
%
Enqueued test cases are stored along with a given mutation budget, that controls
how many times a given test case can be mutated before being finally thrown
away.
%
This budget is calculated using a heuristic derived from AFL's power schedule,
i.e., more budget to test cases that lead to shorter executions, or that
discover more parts of the code.
%
On each iteration, the algorithm selects the next test case by mutating the
first value on the queue of valid test cases or, alternatively, from the queue
of discarded test cases if the former one is empty.
%
If none of the queues contain test cases to mutate, CGPT generates a new value
from scratch.
%
Then, the algorithm run this test case and evaluates whether it was interesting
(i.e., it exercises a new path) based on its trace information.
%
If the test case does in fact discover a new path, it is enqueued at the end of
its corresponding queue, depending on whether it passed the property
precondition or was discarded.
%
%% (i.e., it failed the precondition).
%
This process alternates between generation and mutation until a bug is found or
it reaches the test limit.

%% %
%% Enqueued values are stored along with a given mutation budget, that controls how
%% many times a given test case can be mutated before being finally thrown away.

%% Once inside of the main loop, the algorithm picks the next test case using a
%% simple criterion: if there are valid values enqueued for mutation, it picks the
%% first one, mutates it and returns it, decreasing its energy by one.
%% %
%% If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
%% %
%% If none of the mutation queues contain any candidates, CGPT generates a new
%% value from scratch.
%% %
%% This selection process is illustrated in detail in Algorithm
%% \ref{algo:fuzzchick:pick}.

%% Having selected the next test case, the main loop proceeds to execute it,
%% capturing both the result (passed, discarded due to preconditions, or failed)
%% and its execution trace over the system under test.
%% %
%% If the test case fails, it is immediately reported as a bug.
%% %
%% If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
%% new path) based on its trace information and the one from previously executed
%% test cases (represented by \textit{TLog}).
%% %
%% If the test case does in fact discover a new path, it is enqueued at the end of
%% its corresponding queue, depending on whether it passed or was discarded.
%% %
%% %% (i.e., it failed the precondition).
%% %
%% This process alternates between generation and mutation until a bug is found or
%% we reach the test limit.

%% In addition, high-level, structure-preserving mutators are better suited for
%% producing syntactically valid mutants, avoiding the time wasted by using generic
%% low-level mutators that act over the ``serialized'' test cases and know very
%% little about their structure, thus producing syntactically broken mutants most
%% of the time.
%% %
%% This grammar-aware mutation technique has shown to be quite useful when fuzzing
%% systems accepting structurally complex inputs.
%% %
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%% %
%% In CGPT, external grammars are not required.
%% %
%% The data types used by the inputs of the testing properties already describe the
%% structure of the random data we want to mutate in a concrete manner, and
%% specialized mutators acting at the data constructor level can be automatically
%% derived directly from their definition.

%% \paragraph{Testing Loop}
%% Outlined in Algorithm \ref{algo:fuzzchick:loop}, CGPT starts by creating two
%% queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
%% executed test cases, respectively.
%% %
%% Enqueued values are stored along with a given mutation budget, that controls how
%% many times a given test case can be mutated before being finally thrown away.

%% Once inside of the main loop, the algorithm picks the next test case using a
%% simple criterion: if there are valid values enqueued for mutation, it picks the
%% first one, mutates it and returns it, decreasing its energy by one.
%% %
%% If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
%% %
%% If none of the mutation queues contain any candidates, CGPT generates a new
%% value from scratch.
%% %
%% This selection process is illustrated in detail in Algorithm
%% \ref{algo:fuzzchick:pick}.

%% Having selected the next test case, the main loop proceeds to execute it,
%% capturing both the result (passed, discarded due to preconditions, or failed)
%% and its execution trace over the system under test.
%% %
%% If the test case fails, it is immediately reported as a bug.
%% %
%% If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
%% new path) based on its trace information and the one from previously executed
%% test cases (represented by \textit{TLog}).
%% %
%% If the test case does in fact discover a new path, it is enqueued at the end of
%% its corresponding queue, depending on whether it passed or was discarded.
%% %
%% %% (i.e., it failed the precondition).
%% %
%% This process alternates between generation and mutation until a bug is found or
%% we reach the test limit.

%% \begin{algorithm}
%%   \SetArgSty{textnormal}
%%   %% \SetInd{0em}{0.75em}
%%   \SetKw{KwNot}{not}
%%   \SetAlgoLined
%%   \DontPrintSemicolon
%%   \SetKwFunction{FC}{Loop}
%%   \SetKwProg{Fn}{Function}{:}{}
%%   \Fn{\FC{P, N, gen, mut}}{
%%     i $\gets$ 0\;
%%     TLog, QSucc, QDisc $\gets$ $\varnothing$\;
%%     \While{i $<$ N}{
%%       x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
%%       (result, trace) $\gets$ WithTrace(P(x))\;
%%       \lIf{\KwNot result}{\KwRet Bug(x)}
%%       \If{Interesting(TLog, trace)}{
%%         e $\gets$ Energy(TLog, x, trace)\;
%%         \eIf{\KwNot Discarded(result)}{
%%           Enqueue(QSucc, (x, e))\;
%%         }{
%%           Enqueue(QDisc, (x, e))\;
%%         }
%%       }
%%       i $\gets$ i+1\;
%%     }
%%     \KwRet Ok\;
%%   }
%% \caption{\label{algo:fuzzchick:loop}CGPT Testing Loop}
%% \end{algorithm}

%% \begin{algorithm}
%%   \SetArgSty{textnormal}
%%   %% \SetInd{0em}{0.75em}
%%   \SetKw{KwNot}{not}
%%   \SetAlgoLined
%%   \DontPrintSemicolon
%%   \SetKwFunction{Pick}{Pick}
%%   \SetKwProg{Fn}{Function}{:}{}
%%   \Fn{\Pick{QSucc, QDisc, gen, mut}}{
%%     \If{\KwNot Empty(QSucc)}{
%%       (x,e) $\gets$ Deque(QSucc)\;
%%       \If{e $>$ 0}{
%%         PushFront(QSucc, (x, e-1))
%%       }
%%       \KwRet Sample(mut(x))\;
%%     }
%%     \uElseIf{\KwNot Empty(QDisc)}{
%%       (x,e) $\gets$ Deque(QDisc)\;
%%       \If{e $>$ 0}{
%%         PushFront(QDisc, (x, e-1))
%%       }
%%       \KwRet Sample(mut(x))\;
%%     }
%%     \lElse{\KwRet Sample(gen)}
%%   }
%% \caption{\label{algo:fuzzchick:pick}CGPT Seed Selection}
%% \end{algorithm}

%% %% \subsubsection{Type-preserving mutations}

%% \paragraph{Structure-preserving Mutators}

%% %% are no more than specialized random generators, parameterized
%% %% by the original input to be mutated.
%% %% %
%% %% They

%% Mutators in GCPT use a simple set of mutation operations that are randomly
%% applied to their inputs at the data type level.
%% %
%% In simple terms, these operations encompass
%% %
%% \begin{inparaenum}
%% \item \emph{swapping a data constructor}, replacing its outermost data
%%   constructor by another one, e.g., a \texttt{Leaf} by a \texttt{Branch} and
%%   vice-versa, either dropping unused subexpressions or generating new random
%%   ones when the new data constructor contains extra fields that cannot be filled
%%   using existing ones;
%% %% \item \emph{shrinking the value}, replacing its outermost data constructor with
%% %%   one that contains a subset of its fields, reusing existing subexpressions;
%% %% \item \emph{growing the value}, replacing its outermost data constructor with
%% %%   one that contains a superset of its fields, reusing existing subexpressions
%% %%   and generating random ones when needed;
%% \item \emph{returning a subexpression of the same type}, e.g, returning a left
%%   or right subtree when applied to a \texttt{Branch}; and
%% \item \emph{mutating recursively}, applying a mutation operation over an
%%   immediate subexpression.
%% \end{inparaenum}
%% %
%% We will provide more details about structure-preserving mutations in Section
%% \ref{sec:mutagen}.

@@

\item \emph{Power schedule:}
%
it is not clear how the power schedule used to assign a given budget to each
mutable test case in CGPT works in the context of high-level
structure-preserving mutations.
%
If it assigns too much budget to certain not-so-interesting seeds, some bugs
might not be discovered on a timely basis.
%
Conversely, assigning too little budget to interesting test cases might mean
that some bugs cannot be discovered at all --- randomly generating the same test
case again later does not help, as it becomes uninteresting based on historic
trace information.

@@

%% The next two lines define the bibliography style to be used, and the
%% bibliography file.
\bibliographystyle{ACM-Reference-Format}
%% \bibliographystyle{plainnat}
\bibliography{references}
